\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{ulem}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Drowsiness Detection Project Report}

\author{Dinh Duc Dat Nguyen \\
Oregon State University \\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Feng-Shih Cheng \\
Oregon State University \\
{\tt\small karaburi2023@gmail.com}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top
   of the left-hand column, below the author and affiliation
   information. Use the word ``Abstract'' as the title, in 12-point
   Times, boldface type, centered relative to the column, initially
   capitalized. The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

\section{Introduction}
The introduction should include the research motivation, objectives, and your main contributions.
\begin{itemize}
    \item Introduce the importance of this topic in the field of deep learning.
    \item Briefly describe the current challenges.
    \item Summarize the main work of this paper, for example: ``We evaluated ...'' or ``We implemented ...''.
\end{itemize}

At the pace of modern society continues to accelerate, fatigue due to insufficient sleep of drivers has become one of critical road-safety issues. 
According to data reported by the AAA Foundation for Traffic Safety~\cite{Tefft2016}, crash risk increases greatly as sleep duration decreases. For example, drivers who sleep 6 to 7 hours exhibit a 1.3-fold increase in crash risk relative to fully rested drivers, 
rising to 1.9-fold with only 5-6 hours of sleep.

Between 2017 and 2021, 17.6\% of all ftal crashes involved drowsy driving~\cite{tefft2024drowsy}. 
This issue keeps continuing today; As of Novermber 2025, the Colorado Sate Patrol reported 627 related incidents~\cite{patla2025drowsy}, 
and A 2024 survey indicate that 25 \% of drivers admitted to having engaged in drowsy driving within the past 30 days~\cite{AAA2024TSCI}.

Though car manufacturers such as Tesla and Mercedes-Benz have brought in advanced driver-assistance systems (ADAS) into their vehicles~\cite{abd2023efficient,jabbar2020driver},
accurately assessing a driver's real-time physiological and cognitive state through non-invasive methods remains a significant and unresolved challenge at the field of modern automotive safety.

In order to successfully prevent traffic accidents, detecting drowsiness is a fundamental and valuable requirement. 

%TODO: ADD few related work intro here
傳統的疲勞偵測多仰賴 PERCLOS 等啟發式指標，然而這些方法在複雜光影或個體差異下表現

Therefore, we were looking to develop a program capable of predicting when a driver may become drowsy and immediately notifying them to avoid serious accidents on the road. 
To achieve this goal, we need to track the driver's eye-opening and eye-closing patterns to determine whether the driver is in a drowsy-driving state. 
To address this problem, we applied deep learning methods, which will be explained in the following chapters.

Inspired by Hassan et al.~\cite{Hassan2025}, this project adopts a simpler approach to develop an effective eye-state classification system (open vs.\ closed) based on a deep learning model trained on the MRL dataset, while applying the best-performing model to provide real-time eye-state feedback.
The proposed method consists of a series of steps starting with \dots

% answer following bullet points question and convert them into a paragraph form
\begin{itemize}
    \item Data preparation
    \item Do we change the the size of the image? normalization? data augmentation?
    \item What is our Model?
    \item After training and evaluating? how we pick the best model? (accuracy, precision, recall and F1-score)
\end{itemize}

% Replace this for real structure of the paper.
The rest of the paper is structured as follows: In “Literature review” section, present a critical appraisal of related works. In “Proposed methodology” section, we discuss the methodology proposed. “Experiments and results” section presents the experimental results. “Comparative study” section presents the comprehensive study. In “Discussion” section, the discussion is presented. “Conclusion and future directions” section: concludes this paper.

\section{Background}
Provide the background knowledge needed to understand this study.
\begin{itemize}
    \item Relevant theoretical foundations (e.g., Transformer architectures and the biological significance of TSS prediction).
    \item Problem definition: define the specific mathematical or engineering problem you aim to solve.
\end{itemize}

\section{Related Works}

Several studies have developed real-time eye state recognition systems designed for resouce-constrained embedded devices. 

For example, DriCare represents a great advancement in fatigue detection using video-based facial analysis~\cite{8808931}. To overcome the tracking inaccuracies that previous algorithms suffered from, 
the authors introduced a new method, which enables more robust face tracking. The core of DriCare's improvement lies in detecting 68 major facial characteristics to define facial regions. 
By analyzing temporal cues like yawning frequency, blinking rate and the the duration of eye closure, the system can effectively identify signs of fatigue. 
Experimental results demonstrate that DriCare achieves a high accuracy of approximately 92\%.

To mitigate limitations of visible-light cameras under fluctuating light and vision, an study has introduced faical thermal imaging and infrared-based illuminators ~\cite{kiashari2020evaluation}. 
These approaches evaluate drowsiness by extracting metrics such as Percentage of Eye Closure (PERCLOS), Eye Closing Duration (ECD) and Average Mouth Opening Time (AOT), along with physiological features like inspiratory-to-Expiratory ratio. 
By using fuzzy logic framework and SVM classifiers, the appraoch categorizes fatigue into three levels with up to 91\% accuracy.

More recently, a study using Transformer-based architectures has pushed accuracy even higher, reaching 99.15\% on the MRL Eye Dataset using a Vision Transformer (ViT)~\cite{Hassan2025}. 
That research also evaluated other models like the Swin Transformer and used Class Activation Mapping (CAM) to show which parts of the eye the model focuses on. 
% TO double check!
Rather than adopting their Transformer-based architecture, our implementation uses a lighter technical stack. 
At the same time, our method shares the same objective and methodological approach: inferring driver fatigue from eye-state dynamics, including preprocessing steps such as normalization and an explicit focus on robustness in evaluation. 
This design choice makes the system more practical for real-world deployment on resource-constrained in-vehicle platforms.

\begin{itemize}
    \item Explain how your work differs from or improves upon existing studies.
\end{itemize}

\section{Technical Approach}
In this project, we implement a lightweight convolutional neural network (CNN) specifically optimized for binary eye-state classification. The architecture consists of three convolutional layers designed to extract spatial features from eye images effectively.

\textbf{Neural Network Architecture.} 
The proposed model, \textit{EyeClassifier}, follows a sequential structure:
\begin{itemize}
    \item \textbf{Feature Extraction:} Three convolutional layers with 32, 64, and 128 filters ($3 \times 3$ kernel), each followed by a ReLU activation and a $2 \times 2$ Max-Pooling layer to reduce spatial dimensions.
    \item \textbf{Classification Head:} A fully connected layer with 128 neurons, followed by a Dropout layer ($p=0.5$) to prevent overfitting, and a final output neuron with a Sigmoid activation function to predict the probability of the eye being open.
\end{itemize}

\textbf{Loss Function and Optimizer.} 
Since the task is a binary classification, we employ the Binary Cross Entropy (BCE) loss:
\begin{equation}
L = -\frac{1}{N} \sum_{i=1}^N [y_i \log(\hat{y}_i) + (1 - y_i) \log(1 - \hat{y}_i)]
\end{equation}
where $y_i$ is the ground truth and $\hat{y}_i$ is the predicted probability. We use the Adam optimizer for its efficient gradient descent properties.

\section{Experiments}
\subsection{Experimental Design}

The experiments are conducted using the MRL Eye Dataset~\cite{}, which is a large-scale collection of eye images specifically designed for drowsiness detection. 
The dataset contains 84,898 images captured under various conditions, including different lighting environments, gender distributions and the presence of eyeglasses. 
Each image in the dataset follows a specific naming convention, from which we extract the eye-state label (0 for closed, 1 for open) as the ground truth for our binary classification task.

\textbf{Dataset.} 
We utilize the MRL Eye Dataset, which contains 84,898 grayscale images. For our experiments, all images are resized to $24 \times 24$ pixels and normalized to the range $[-1, 1]$. We ensure a robust evaluation by splitting the data into training and testing sets.

\textbf{Experimental Setup.} 
The model is implemented in Python using the PyTorch library. Training is conducted on a macOS environment utilizing the Metal Performance Shaders (MPS) for hardware acceleration. The training process is configured with the following hyperparameters:
\begin{itemize}
    \item \textbf{Batch Size:} 64 samples per iteration.
    \item \textbf{Epochs:} 5 iterations over the entire dataset.
    \item \textbf{Learning Rate:} 0.001.
\end{itemize}

\textbf{Evaluation Metrics.} 
To quantitatively assess the model, we use Accuracy, Precision, Recall, and the F1-score. These metrics allow us to evaluate not only the overall performance but also the reliability of the system in detecting critical ``eye-closed'' states that indicate driver drowsiness.


\subsection{Results and Discussion}
Present your results.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Accuracy \\
\hline\hline
Baseline & 85\% \\
Our Method & 92\% \\
\hline
\end{tabular}
\end{center}
\caption{Performance comparison of different methods.}
\end{table}

\section{Conclusion and Future Work}
Summarize your research findings, and discuss the current limitations as well as possible future improvements.


{\small
\bibliographystyle{ieee}
\bibliography{dp_final}
}

\end{document}
